### 2017-03-20 paper/sections/0_abstract.tex
paper: submit abstract and touch up intro
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}

HPC file systems are abandoning POSIX because the synchronization and
serialization overheads are too costly -- and often uneccessary -- for their
applications.  We provide an API for the client to (1) merge updates into the
global namespace and (2) assign policies to subtrees so that other clients
cannot interfere with the decoupled namespace.  This allows administrators to
optimize subtrees within the same namespace for different workloads. We draw
conclusions about the performance impact of previously unexplored
consistency/durability metadata designs and show that consistency can cause a
104\(\times\) slow down while merging updates (7\(\times\) slow down) and
maintaining durability (\(10\times\) slow down) have a more reasonable cost.

\end{abstract}