### 2017-03-29 paper/sections/0_abstract.tex
paper: wordsmith and push related work backwards
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\begin{abstract}

HPC file systems are abandoning POSIX because the synchronization and
serialization overheads are too costly -- and often uneccessary -- for their
applications.  We provide an API for the client to (1) decouple subtrees from
the global namespace, (2) merge metadata updates back into the global namespace
and (3) assign policies to subtrees so that other clients cannot interfere with
the decoupled namespace.  This allows administrators to optimize subtrees
within the same namespace for different workloads. We draw conclusions about
the performance impact of previously unexplored consistency/durability metadata
designs and show that consistency can cause a 104\(\times\) slow down while
merging updates (7\(\times\) slow down) and maintaining durability
(\(10\times\) slow down) have a more reasonable cost.

\end{abstract}